---
title: "mazeR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mazeR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center"
)
library(tidyr)
library(ggplot2)
library(mazeR)
set.seed(666)

```

The mazeR package is based on the [randomized depth-first search](https://en.wikipedia.org/wiki/Maze_generation_algorithm) maze generation algorithm.

A Maze object can be created by calling the `generate_maze` function. The Maze can then be displayed using the `render_maze` function.  

```{r}
maze_a <- generate_maze(25, 25)
render_maze(maze_a)
```

The option to display the maze solution can be enabled using the `show_solution` argument of `render_maze`.

```{r}
render_maze(maze_a, show_solution = TRUE)
```

The default start and exit locations of the maze are the lower left and upper right coordinates, respectively. The exit can be manually specified using the `exit` argument.

```{r}
maze_b <- generate_maze(25, 25, exit = c(12, 20, 1))
render_maze(maze_b)
```

Some arguments have the ability to influence the randomness of the search algorithm. The `i_weight`, `j_weight`, and `k_weight` arguments modify the probabilities to favor vertical, horizontal, or normal path searches, respectively. For example, setting `i_weight = 2` favors vertical movement 2x more that usual.

```{r}
maze_c <- generate_maze(25, 25, i_weight = 2)
render_maze(maze_c)
```

Setting `i_weight = 10` exaggerates this effect even more.

```{r}
maze_d <- generate_maze(25, 25, i_weight = 10)
render_maze(maze_d)
```

The `j_weight` argument influences the tendency for more horizontal corridors. 

```{r}
maze_e <- generate_maze(25, 25, j_weight = 5)
render_maze(maze_e)
```

The `curl*` arguments can also be used to influence the path through the space. `curl` represents an angle between the path parallel to the most recent direction of travel and a vector of length `curl_weight`, indicating the direction and magnitude of the curl effect. The parallel components of this vector are then added to each "potential next direction" vector, resulting in revised movement probabilities. Positive curl corresponds to a tendency toward clockwise rotation.

```{r echo=FALSE}

curl_ex <- 0.3

p <- 
ggplot() +
  annotate("rect", xmin = -1.5, ymin = -0.5, xmax = -0.5, ymax = 0.5, fill = "springgreen", alpha = 0.5) +
  annotate("rect", xmin = -0.5, ymin = 0.5, xmax = 0.5, ymax = 1.5, fill = "springgreen", alpha = 0.5) +
  annotate("rect", xmin = 0.5, ymin = -0.5, xmax = 1.5, ymax = 0.5, fill = "springgreen", alpha = 0.5) +
  annotate("rect", xmin = -0.5, ymin = -1.5, xmax = 0.5, ymax = -0.5, fill = "gray80", alpha = 0.5) +
  annotate("text", x = 0, y = -1.3, label = "Last cell", size = 3) +
  annotate("rect", xmin = -0.5, ymin = -0.5, xmax = 0.5, ymax = 0.5, fill = "dodgerblue", alpha = 0.5) +
  annotate("text", x = 0, y = -0.3, label = "Current cell", size = 3) +
  geom_blank(data = data.frame(x = c(-2, 2), y = c(-2, 2)), aes(x = x, y = y)) +
  coord_equal() +
  scale_x_continuous() +
  scale_y_continuous() +
  theme(panel.background = element_rect(fill = NA),
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank())

curl1 <- 
p + 
  geom_segment(aes(x = 0, y = 0, xend = 1, yend = 0), lwd = 1, 
               arrow = arrow(length = unit(0.3, "cm"), type = "closed"), lineend = "round", linejoin = "mitre") +
  geom_segment(aes(x = 0, y = 0, xend = -1, yend = 0), lwd = 1, 
               arrow = arrow(length = unit(0.3, "cm"), type = "closed"), lineend = "round", linejoin = "mitre") +
  geom_segment(aes(x = 0, y = 0, xend = 0, yend = 1), lwd = 1, 
               arrow = arrow(length = unit(0.3, "cm"), type = "closed"), lineend = "round", linejoin = "mitre") +
  annotate("text", x = -1, y = 0.3, label = "Possible next cell", size = 3) +
  annotate("text", x = 0, y = 1.3, label = "Possible next cell", size = 3) +
  annotate("text", x = 1, y = 0.3, label = "Possible next cell", size = 3) +
  
  geom_point(aes(x = 0, y = 0), size = 5) +
  ggtitle("Possible next moves")

curl2 <- 
p +
  geom_segment(aes(x = 0, y = 0, xend = sin(pi / 2 * curl_ex), yend = cos(pi / 2 * curl_ex)), lwd = 1, 
               arrow = arrow(length = unit(0.3, "cm"), type = "closed"), lineend = "round", linejoin = "mitre") +
  # geom_segment(aes(x = 0, y = 0, xend = 0, yend = 1), lty = 2, color = "gray60") +
  annotate("text", x = 0.1 * sin(pi / 2 * curl_ex) + 0.1, y = 0.1 * cos(pi / 2 * curl_ex), angle = 90 - 27, label = "curl_weight", size = 3, hjust = 0, vjust = 0.5) +
  geom_curve(aes(x = 0, y = 0.3, xend = 0.3 * sin(pi / 2 * curl_ex), yend = 0.3 * cos(pi / 2 * curl_ex)), curvature = -0.3) +
  annotate("text", x = -0.05, y = 0.4, label = "\u03C0/2*curl", size = 3, hjust = 1, vjust = 0.5) +
  geom_segment(aes(x = 0, y = 0, xend = 1, yend = 0), lwd = 1, 
               arrow = arrow(length = unit(0.3, "cm"), type = "closed"), lineend = "round", linejoin = "mitre", color = "gray60") +
  geom_segment(aes(x = 0, y = 0, xend = -1, yend = 0), lwd = 1, 
               arrow = arrow(length = unit(0.3, "cm"), type = "closed"), lineend = "round", linejoin = "mitre", color = "gray60") +
  geom_segment(aes(x = 0, y = 0, xend = 0, yend = 1), lwd = 1, 
               arrow = arrow(length = unit(0.3, "cm"), type = "closed"), lineend = "round", linejoin = "mitre", color = "gray60") +
  
  geom_point(aes(x = 0, y = 0), size = 5) +
  ggtitle("Influence of curl")

curl3 <-  
p + 
  geom_segment(aes(x = 0, y = 0, xend = 1, yend = 0), lwd = 1, color = "gray60",
               arrow = arrow(length = unit(0.3, "cm"), type = "closed"), lineend = "round", linejoin = "mitre") +
  geom_segment(aes(x = 0, y = 0, xend = -1, yend = 0), lwd = 1, color = "gray60", 
               arrow = arrow(length = unit(0.3, "cm"), type = "closed"), lineend = "round", linejoin = "mitre") +
  geom_segment(aes(x = 0, y = 0, xend = 0, yend = 1), lwd = 1, color = "gray60", 
               arrow = arrow(length = unit(0.3, "cm"), type = "closed"), lineend = "round", linejoin = "mitre") +
  geom_segment(aes(x = 0, y = 0, xend = sin(pi / 2 * curl_ex), yend = cos(pi / 2 * curl_ex)), lwd = 1, color = "gray60", 
               arrow = arrow(length = unit(0.3, "cm"), type = "closed"), lineend = "round", linejoin = "mitre") +
  geom_segment(aes(x = 0, y = 0, xend = 1 + sin(pi / 2 * curl_ex), yend = 0), lwd = 1, color = "purple", 
               arrow = arrow(length = unit(0.3, "cm"), type = "closed"), lineend = "round", linejoin = "mitre") +
  geom_segment(aes(x = 0, y = 0, xend = -1 + sin(pi / 2 * curl_ex), yend = 0), lwd = 1, color = "purple", 
               arrow = arrow(length = unit(0.3, "cm"), type = "closed"), lineend = "round", linejoin = "mitre") +
  geom_segment(aes(x = 0, y = 0, xend = 0, yend = 1 + cos(pi / 2 * curl_ex)), lwd = 1, color = "purple", 
               arrow = arrow(length = unit(0.3, "cm"), type = "closed"), lineend = "round", linejoin = "mitre") +
  
  geom_point(aes(x = 0, y = 0), size = 5) +
  ggtitle("Revised next moves")


```

```{r fig.width=8, fig.height=3, echo=FALSE}

gridExtra::grid.arrange(curl1, curl2, curl3, nrow = 1)

```


```{r}
maze_f <- generate_maze(25, 25, curl = 0.3, curl_prob = 1.0, curl_weight = 2, exit = c(12, 12, 1))
render_maze(maze_f)
```

The curl effect is easier to see when the solution is shown.

```{r}
render_maze(maze_f, show_solution = TRUE)
```

`curl = 0` increases the tendency to continue in straight lines.

```{r}
maze_g <- generate_maze(25, 25, curl = 0, curl_prob = 1.0, curl_weight = 10, exit = c(12, 12, 1))
render_maze(maze_g)
```


```{r}

maze_h <- generate_maze(25, 25, curl = 1, curl_prob = 0.1, curl_weight = 10, uncurl = TRUE)
render_maze(maze_h, show_solution = TRUE)

```

